---
title: "Homework 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Analysis of Lecture 3 data
=========================
Oct 15, 2018

# Learn to manupulate Tidy Data via four new functions
This document will explain how to use four new functions in the Tidyverse 
package: mutate, join functions, summarize, group_by

## Load Tidyverse 
Run `library(tidverse)` to load the tidyverse library into RStudio

```{r}
library(tidyverse)
```

## Read in data file
This is pulling the data file from a GitHub repository in the data directory

```{r}
data <- read_tsv("../data/example_dataset_2.tsv") %>% 
  print()
```

## Mutate Function

### Use mutate to create new columns
`mean_ratio` is the title of your new column

```{r}
data <- data %>%
  mutate(mean_ratio = mean_yfp / mean_rfp) %>%
  print()
```

### Use mutate to modify existing columns
This set of code rounds the mean_ratio column to 2 digits
NOTE: R has functions for almost all mathematical and statistical manipulation

```{r}
data %>%
  mutate(mean_ratio = round(mean_ratio, 2)) %>%
  print()
```

## TYPE_join functions

### Read in annotations file
This second table of data is given the name annotations for reference in future 
code

```{r}
annotations <- read_tsv("../data/example_dataset_3.tsv") %>%
  print()
```

### inner_join keeps common rows
Keep rows that are present in both and then it aligns the new columns
NOTE: Don’t need to have the same number of rows in each data set
```{r}
data %>%
  inner_join(annotations, by = "strain") %>%
  print()
```

### left_join keeps all rows in left data frame
Takes data frame on the right and joins it with data frame on the left
Throws out rows only in the right data frame (annotations)

```{r}
data %>%
  left_join(annotations, by = "strain") %>%
  print()
```

### right_join keeps all rows in right data frame
Takes data frame on the left and joins it with data frame on the right
Throws out rows only in the left data frame (data)

```{r}
data %>%
  right_join(annotations, by = "strain") %>%
  print()
```

## Summarize Functions

### Use sammarize to calculate stats across rows
This outputs the max value of only what you ask

```{r}
data %>%
  summarize(max_yfp = max(mean_yfp),
            max_rfp = max(mean_rfp)) %>%
  print()
```

Other examples of summary functions:
min() mean() sd() first() n()

## Group_by Function

### Read in new file
This file includes the replicate of each strain

```{r}
data <- read_tsv("../data/example_dataset_4.tsv") %>%
  print(n = 10)
```

### Use group_by to group subsets of rows
Subsets rows by the column you are interested in
- Example: group by strain

```{r}
data %>%
  group_by(strain) %>%
  print(n = 10)
```

### group_by + summarize for statistics by group
This will apply the summarize function to each one of the groups

```{r}
data %>%
  group_by(strain) %>%
  summarize(mean_yfp = mean(yfp), mean_rfp = mean(rfp)) %>%
  print()
```

Add a standard error column by group

```{r}
data %>%
  group_by(strain) %>%
  summarize(mean_yfp = mean(yfp), 
            mean_rfp = mean(rfp), 
            se_yfp = sd(yfp) / sqrt(n()),
            se_rfp = sd(rfp) / sqrt(n())) %>%
  print()
```

## All Functions Combined

### %>% enables complex data analysis piplines
Makes an easy way to read how the data is combined and manipulated

```{r}
data %>%
  group_by(strain) %>%
  summarize(mean_yfp = mean(yfp), mean_rfp = mean(rfp)) %>%
  mutate(mean_ratio = mean_yfp / mean_rfp) %>%
  left_join(annotations, by = "strain") %>%
  print()
```

### %>% and + if you want to plot

```{r}
data %>%
  group_by(strain) %>%
  summarize(mean_yfp = mean(yfp), mean_rfp = mean(rfp)) %>%
  mutate(mean_ratio = mean_yfp / mean_rfp) %>%
  left_join(annotations, by = "strain") %>%
  ggplot(aes(x = kozak_region, y = mean_ratio,
             color = insert_sequence, group = insert_sequence)) +
  geom_line() +
  geom_point()
```

Now you can use all of these Tidyverse functions on your own.
Good Luck